---
title: CELLO-Seq Sarlacc pipeline - simulated UMI grouping
author: Rebecca Berrens, Andrian Yang, Aaron Lun, Florian Bieberich
output:
  BiocStyle::html_document:
    toc_float: true
    titlecaps: false
---

```{r, echo=FALSE, results="hide", message=FALSE}
require(knitr)
opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

```{r setup, echo=FALSE, message=FALSE}
library(sarlacc)
```

# Introduction

This series of Rmarkdown files forms the CELLO-Seq sarlacc data processing pipeline, which utilises the `sarlacc` R package as its base.

For perfect grouping of reads, we will create read groups based on the read names of the reads, which contain the true location of the reads.

# Reading files

We read in the the RDS file containing the adapter information for the sample.

```{r}
my.sample_fastq <- readRDS(paste0(INDEX, ".rds"))
```

# Defining perfect groups

We now form the perfect groups based on the read names.

```{r}
sample.readnames <- gsub(",.*", "", rownames(my.sample_fastq))
groups <- lapply(unique(sample.readnames), function(x) which(sample.readnames == x))
summary(lengths(groups))
```

# Writing out read groups

We will now output the read groups for error-correction or deduplication.
We will also output the read names for each read groups since the read names contains the true location of each simulated reads.
Since error-correction can take some time to run, we will split the list of read groups into multiple smaller chunks and limit the amount of reads per groups.

```{r}
read.split <- 2000000
max.read.group.size <- 50

chunk.id <- 1
chunk.groups <- c()
chunk.read.total <- 0

for (i in 1:length(groups)) {
  group <- groups[i]
  group.length <- length(unlist(group))

  if (group.length == 0){
    next
  }

  if (group.length > max.read.group.size) {
    group[[1]] <- sample(group[[1]], max.read.group.size)
    group.length <- max.read.group.size
  }

  if (length(chunk.groups) == 0) {
    chunk.groups <- c(chunk.groups, c(group))
    chunk.read.total <- chunk.read.total + group.length
  } else {
    if ((chunk.read.total + group.length) > read.split) {
      saveRDS(chunk.groups, file=paste0(INDEX, ".perfect_groups.", chunk.id, ".rds", sep=""))

      read.names = lapply(chunk.groups, function(x){
        rownames(my.sample_fastq[x,])
      })
      saveRDS(read.names, paste0(INDEX, ".perfect_readnames.", chunk_id, ".rds", sep=""))

      chunk.id <- chunk.id + 1
      chunk.groups <- c()
      chunk.read.total <- 0
    }

    chunk.groups <- c(chunk.groups, c(group))
    chunk.read.total <- chunk.read.total + group.length
  }
}
saveRDS(chunk.groups, file=paste0(INDEX, ".perfect_readnames.", chunk.id, ".rds", sep=""))

read.names = lapply(chunk.groups, function(x){
  rownames(my.sample_fastq[x,])
})
saveRDS(read.names, paste0(INDEX, ".perfect_readnames.", chunk_id, ".rds", sep=""))
```

# Session information

```{r}
sessionInfo()
```




